Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-27 paper_jasa.tex
summary table heading
power comparison discussion

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. }
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Total &Lineup\\ 
    \hline
  1 & 239 & 121 & 107 & 2317 & 60 \\ 
  2 & 351 & 185 & 164 & 3858 & 70 \\ 
  3 & 156 & 106 &  50 & 1681 & 30 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

Results for experiment 3, are quite different. This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes. There is no actual conventional power in this situation, because assumptions are violated. It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model. The crosses represent \green{... please fill this in}. It is really curious to see that the power of the visual test is really high when effect is small. These results are based on whether the subject correctly picked the actual data plot, regardless of reason. Although they were asked to select the plot that exhibited thr highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data. This will be explored further in Section \ref{TypeIII}.

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
       \caption{Power in comparison to effect for the three experiments. Points indicate subject responses, with size indicating count. Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup. The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.  Conventional test power is drawn as a dashed line. For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data. \green{These next two sentences are confusing. Maybe too much for caption, and should be moved to text.} Crosses mark points that are the conventional power of true effect of the contaminated data. We obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated and plotted as cross marked points. Notice that the power structure is completely broken for contaminated data. For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3.
}
       \label{fig:power_loess_effect}
\end{figure}

@@

\subsection{Alternative - Type III error}~\label{TypeIII}

>>>>

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. }
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
  1 & 239 & 121 & 107 & 2317 & 60 \\ 
  2 & 351 & 185 & 164 & 3858 & 70 \\ 
  3 & 156 & 106 &  50 & 1681 & 30 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

Results for experiment 3, are quite different. This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes. There is no actual conventional power in this situation, because assumptions are violated. It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model. The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks. For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated. Notice that the power structure is completely broken for contaminated data. For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3

It is really curious to see that the power of the visual test is really high when effect is small. These results are based on whether the subject correctly picked the actual data plot, regardless of reason. Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data. This will be explored further in Section \ref{sec:TypeIII}.

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
       \caption{Power in comparison to effect for the three experiments. Points indicate subject responses, with size indicating count. Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup. The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.  Conventional test power is drawn as a dashed line. For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
}
       \label{fig:power_loess_effect}
\end{figure}

@@

\subsection{Alternative - Type III error}~\label{sec:TypeIII}
